---
title: "Day 2. Session 3"
author: "Hector Najera"
date: "2025-06-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(descr)
library(flextable)
```

```{r}
health_data<-read_excel("health_dataexample.xlsx")
```

Let's explore the data

```{r}
str(health_data)
```

What are the names of the variables in my data set?

```{r}
names(health_data)
```

Exploring the variables

```{r}
summary(health_data$age)
```

How would you explore the blood_pressure variable and what can you tell about it?

```{r}
freq(health_data$sex)
```

Now let's see the region variable

```{r}
freq(health_data$region_id)
```

How would you explore the diagnosis variable and what can you tell about it?

# Data cleaning

```{r echo = TRUE}
health_data_clean1 <- health_data %>% filter(age >= 0 & age <= 100)
age_summary <- summary(health_data_clean1$age)

# Convert to data frame
age_summary_df <- data.frame(
  Statistic = names(age_summary),
  Value = as.numeric(age_summary)
)

# Create flextable
flextable(age_summary_df) %>%
  autofit() %>%
  set_caption("Summary Statistics for Age (Valid Range 0–100)")
```

As an alternative, the code below recodes implausible values in the `age` variable by replacing them with missing values (`NA`), and then summarises the cleaned variable:

```{r echo = TRUE}
health_data_clean2 <- health_data %>% mutate(age = ifelse(age < 0 | age > 100, NA, age))
```

```{r echo = TRUE}
# Example: median imputation
median_age <- median(health_data$age[health_data$age >= 0 & health_data$age <= 100], na.rm = TRUE)
```

```{r}
health_data_clean3 <- health_data %>%
  mutate(
    age = ifelse(age < 0 | age > 100, NA, age),
    age = ifelse(is.na(age), median_age, age)
  )

# Summarize the cleaned and imputed age
age_summary3 <- summary(health_data_clean3$age)

# Convert to data frame
age_summary3_df <- data.frame(
  Statistic = names(age_summary3),
  Value = as.numeric(age_summary3)
)

# Create flextable
flextable(age_summary3_df) %>%
  autofit() %>%
  set_caption("Summary Statistics for Age (Invalid and Missing Replaced with Median)")
```


#### Resolving coding inconsistencies

This recoding step addresses inconsistencies in how sex is recorded in the dataset. It maps different codes to unified labels:

 - "M" and "1" are recoded as "Male"

- "F" and "2" are recoded as "Female"

- Any other value, including missing or unrecognized entries, is labeled as "Unknown"

The result is stored in a new variable, sex_clean, allowing you to retain the original values for auditing purposes. 

```{r echo = TRUE}
health_data <- health_data %>%
  mutate(sex_clean = case_when(
    sex %in% c("M", "1") ~ "Male",
    sex %in% c("F", "2") ~ "Female",
    TRUE ~ "Unknown"
  ))
```

```{r}
sex_table <- as.data.frame(table(health_data$sex_clean))
colnames(sex_table) <- c("Sex (Cleaned)", "Count")

# Create flextable
flextable(sex_table) %>%
  autofit() %>%
  set_caption("Frequency of Cleaned Sex Categories")
```
