
# Day 2 Session 3: How to import data into R 

## Libraries  

```{r}
library(tidyverse)
library(readxl)
library(data.table)
library (haven)
library(writexl)
library(sf)
```

## Import data

Importing: (Excel) xlsx
```{r}
E<- read_xlsx("Data1_fake_data_with_errors.xlsx")
```
Importing: (Excel-separated comma) csv
Note: This file needs to be created from excel as it was not provided

```{r}
C<- read.csv("Data1_fake_data_with_errors.csv") 
```

Importing: (Stata) dta
```{r}
S<-read_dta("Cover.dta") 
```

Loading an R file
```{r}
load ("C1")
```

### Adavanced users: Spatial data

shape files (spatial data set)
Note: The (sf) library needs to be installed and loaded
```{r}
M<- st_read("00ent.shp") 
```
```{r}
plot(M)
```

## Modifying the file by creating a variable

Please create a ID variable for each person in the dataset

```{r}
E$id<- c(1:30)
```

```{r}
names(E)
```


```{r}
str(E$age)
str(E$location)
```


###  Advanced users 

Create a variable in spatial object

```{r}
M$id<- c(1)
```

## Let save the data frame you just work on

save as an R file
```{r}
save(E, file = "E")
```


save in .csv

```{r}
write.csv(E,"E.csv")
```

save in .xlsx
```{r}
write_xlsx(E, "E.xlsx")
```


save in .dta
```{r}
write_dta(E, "E.dta")
```

### Advanced users

save in .shp
```{r}
st_write(M, "M.shp", append=FALSE)
```





