
# Day 3- Session 2: Using  HIES 2023

## Part 1: Prepare data for analysis


```{r echo=F}
library(data.table)
library(haven)
library(dplyr)
library(labelled)
library(descr)
```


### Import data sets

Health data is in stata format 

```{r}
health <- read_dta("p3_health.dta")
```

We also are going to need another data set that contains the island group
```{r}
cover<- read_dta("Cover.dta")
```


### The nesxt step is to select the variables that we want to use for this exercise, which include

* id = "interview__key","hm_basic__id"
* sex= sex
* age= age
* hypertension= p304__8
* island group=  island_group


### Keep the variables that we need

First we work with the health dataset

```{r}
H<- health %>% select(interview__key, sex,age, p304__8)
```


```{r}
I<- cover %>% select(interview__key,island_group)
```



### Merge datasets
Now we need to have all the variables in the same data frame
```{r}
HI<-merge(H,I,by=c("interview__key"), sort=F)
```


## Part 2: Let's conduct some basic analysis

### Calculate the median age 

```{r}
summary(HI$age)
```


### Proportion of male and females
```{r}
HI %>%
  filter(!is.na(sex)) %>%
  count(sex) %>%
  mutate(proportion = n / sum(n)*100)
```


### Proportion of people by island 
```{r}

freq(as_factor(HI$island_group))

HI %>%
  filter(!is.na(island_group)) %>%
  count (as_factor(island_group))%>%
  mutate(proportion = n / sum(n)*100)
```



Recode age groups

* Infants (0-4)
* Children (5-9)
* Adolescents (10-19)
* Adults (20-64)"
* Elderly (65+)


```{r}
HI <- HI %>%
  mutate(age_group = case_when(
    age < 5 ~ "Infants (0-4)",
    age >= 5 & age<= 9 ~ "Children (5-9)",
    age >= 10 & age <= 19 ~ "Adolescents (10-19)",
    age>= 20 & age< 65 ~ "Adults (20-64)",
    age>= 65 ~ "Elderly (65+)"
  ))
```


Calculate the proportion of people with hypertension 

```{r}
 HI %>%
  summarise(prevalence = mean(p304__8, na.rm = TRUE) * 100)
```

## Advanced users


### Crosstabulation between hypertension and sex

```{r}
crosstab(HI$p304__8, HI$sex,prop.c = T,format="SPSS")
```

